### 2014-12-05 docs/ms.Rmd
Editing the intro & methods outline. Begin tightening the table discussion

# Reproducible research in the study of colour in nature

@@

Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data. The rapid rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. As a consequence, the subtle complexity of many analytical techniques [e.g. the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.

Replication and transparency lie at the heart of the scientific enterprise. The full replication of individual studies may often be infeasible @peng_reproducible_2011, but reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the simple practice of complete methods reporting and public data archiving is of tremendous value. It ensures greater comparability between studies, allows for the the independent verification of results, and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].

Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration. We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also assess the availability of publicly archived data and code, and suggest some tools that may be useful for increasing the reproducibility of colour trait research.

## Measuring and analysing colour

@@

Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge. The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966]. Now relatively inexpensive and convenient, spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field. However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture. The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all contribute to variation in data and potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008]. Experience, guidance, consistency and objectivity (e.g. double-blind experimental design) are necessary to ensure data are repeatable and representative of the subject. Crucially, to ensure repeatability across studies, the conditions under which data were collected must be reported exactly.

@@

To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset anonymous as our aim is to make a general statement about methods reporting in the field.

@@

todo: expand table discussion

We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For analytical techniques we characterized a set of criteria required for reproducing the analysis (Table 2). Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results. Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.

@@

Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data. The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], though this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010]. The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods. This is particularly relevant to the field of colour-science, where new methods for analysing colour data are regularly developed [e.g. @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014]. The provision of open data may also foster collaborations as researchers draw upon existing work, and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007]. Given the increasing rate of retractions [@steen_retractions_2010; @van_trouble_2011], there is also impetus for researchers to maintain credibility through transparency. We thus encourage researchers to publicly archive data in as raw a form as possible (e.g. individual reflectance spectra). Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data is relatively straightforward [@whitlock_data_2011]. The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [@neylon_science_2012].

None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.

@@

The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable.

>>>>

# Reproducible research in biological colour-science

@@

Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data. This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. As a consequence, the subtle complexity of many analytical techniques [e.g. the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.

Replication and transparency lie at the heart of the scientific enterprise. While the full replication of individual studies may often be infeasible [@peng_reproducible_2011], the reproductioin of a study's findings by independent researchers is a readily attainable compromise [@peng_reproducible_2006; @peng_reproducible_2011]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the simple practice of complete methods reporting and public data archiving is of tremendous value. It ensures greater comparability between studies, allows for the the independent verification of results, and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].

Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration, and the way in which.... We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research.

## Researcher degrees of freedom in colour analysis

@@

Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge. The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966]. Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as a standard in the field [@andersson_quantifying_2006].

However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture. The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all contribute to variation in data and potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008]. Experience, guidance, consistency and objectivity (e.g. double-blind experimental design) are necessary to ensure data are repeatable and representative of the subject. Crucially, to ensure repeatability across studies, the conditions under which data were collected must be reported exactly.

@@

To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset anonymous as our aim is to make a general statement about methods reporting in the field. Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.

@@

The simplest and most fundamental step in ensuring research is reproducible is through the comprehensive reporting of methods. The myriad researcher degrees of freedom in the capture and analyses of colour data (as outlined above) heighten this need. Accordingly, we developed a list of essential information pertaining to the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility. Given the current state of the field, we focus on spectrometry and photography for the capture of data [as in @kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. Analytical techniques are diverse [kelber_spectral_2010], so we focus on two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling (Table 2). Extra little bit about vis models in particular. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results.

@@

Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data. The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014]. There is evidence that this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010]. The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods. This is particularly relevant to the field of colour-science, where new methods for analysing colour data are frequently developed [e.g. @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014]. The provision of open data may also foster collaborations as researchers draw upon existing work, and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007]. Given the increasing rate of retractions [@steen_retractions_2010; @van_trouble_2011], there is also impetus for researchers to maintain credibility through transparency. We thus encourage researchers to publicly archive data in as raw a form as possible (e.g. individual reflectance spectra). Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data for publication is relatively straightforward [@whitlock_data_2011]. The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [@neylon_science_2012].

(PARA) None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.

@@

The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable....
